---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Title:     "Reproducible Research: Peer Assessment 1"
##            "A Coursera Course"
## Lecturer:  Roger D. Peng, PhD
## Author:    Yosep Thie
<br>
## Loading and preprocessing the data
#### "act" = Activity Data Frame
```{r Reading File, echo=TRUE}
setwd("C:/Users/Ythie/Documents/Github/RepData_PeerAssessment1")
act <- read.csv('activity.csv', header = T)
```
<br>
## What is mean total number of steps taken per day?
#### "acttot" = total number of steps taken per day
#### "actavg" = mean of total number of steps taken per day
#### "actmed" = median of total number steps taken per day
```{r Calculate total number of steps taken per day, and its average and median,      echo=TRUE}
acttot <- aggregate(act$steps ~ act$date, act, sum, na.action = na.pass)
hist(acttot$`act$steps`, xlab = 'Total Number of Steps per Day', main = 
    'Histogram of Total Number of Steps per Day', col = 'Red')

actavg <- aggregate(acttot$`act$steps` ~ acttot$`act$date`, acttot, mean,
                    na.action = na.pass)
head(actavg)
actmed <- aggregate(acttot$`act$steps` ~ acttot$`act$date`, acttot, median,
                    na.action = na.pass)
head(actmed)
```
<br>
## What is the average daily activity pattern?
#### "actint"  = average number of steps taken across all days
#### "maxstep" = interval with maximum number of steps
```{r Plot average daily activity pattern and find interval with maximum number      of steps, echo=TRUE}
actint <- aggregate(act$steps ~ act$interval, act, mean, na.action = na.omit)
plot(actint$`act$interval`, actint$`act$steps`, type = 'l', 
     xlab = '5 Minute Interval', ylab = 'Average Number of Steps',
     main = 'Average Daily Activity Pattern')

maxstep <- subset(actint, actint$'act$steps' == max(actint$'act$steps'))
print('Interval with Maximum Number of Steps = ')
print(maxstep$`act$interval`)
```
<br>
## Imputing missing values
#### "actna"   = activity with missing value(s)
#### "actmean" = average of steps taken per day
#### Imputing Strategy:
####  Missing values of "actmean" are imputed with mean of average steps per day
#### "actmer"  = new data frame by filling out missing values with mean average
#### "acttot2" = new total number of steps taken per day
#### New Histogram displayed from data frame with no missing values
#### "actavg2" = new mean of total number of steps taken per day
#### "actmed2" = new median of total number of steps taken per day
#### Result: HIstogram looks similar but with higher frequency
```{r Impute missing values, echo=TRUE}
actna <- act[!complete.cases(act),]
print('Total Number of Missing Value = ')
dim(actna)

print('Imputing Strategy: Replace each missing value with mean of total average steps taken per day')
actmean <- aggregate(act$steps ~ act$date, act, mean, na.action = na.pass)
actmean$`act$steps`[is.na(actmean$`act$steps`)] <- mean(actmean$`act$steps`,na.rm = T)
names(actmean) <- c('date', 'steps')
actmer <- merge(act, actmean, by = 'date')
actmer$steps.x[is.na(actmer$steps.x)] <- actmer$steps.y[is.na(actmer$steps.x)]
names(actmer) <- c('date', 'steps', 'interval')
actmer <- actmer[,c('date', 'steps', 'interval')]
acttot2 <- aggregate(actmer$steps ~ actmer$date, actmer, sum)
hist(acttot2$`actmer$steps`, xlab = 'Total Number of Steps per Day',
     main = 'Histogram of Total Number of Steps per Day', col = 'Blue')
actavg2 <- aggregate(acttot2$`actmer$steps` ~ acttot2$`actmer$date`, acttot2, mean)
head(actavg2)
actmed2 <- aggregate(acttot2$`actmer$steps` ~ acttot2$`actmer$date`, acttot2, median)
head(actmed2)
```
<br>
## Are there differences in activity patterns between weekdays and weekends?
#### "day"   = factor variable with "Weekday" and "Weekend" values
#### "actwd" = weekday activity subset
#### "actwe" = weekend activity subset
#### "actintwd" = average steps taken across all weekdays
#### "actintwe" = average steps taken across all weekend
#### Plot using base plot
```{r Group daily activity patterns by weekdays and weekends, echo=TRUE}
day <- weekdays(as.Date(act$date))
day[!grepl('^S', day)] <- 'Weekday'
day[grepl('^S', day)] <- 'Weekend'
day <- factor(day)
actmer$day <- day
actwd <- subset(actmer, day == 'Weekday')
actwe <- subset(actmer, day == 'Weekend')
actintwd <- aggregate(actwd$steps ~ actwd$interval, actwd, mean)
actintwe <- aggregate(actwe$steps ~ actwe$interval, actwe, mean)
par(mfrow = c(1,2))
plot(actintwd$`actwd$interval`, actintwd$`actwd$steps`, type = 'l',
     xlab = '5 Minute Interval', ylab = 'Average Number of Steps',
     main = 'Average Weekday Activity Pattern')
plot(actintwe$`actwe$interval`, actintwe$`actwe$steps`, type = 'l',
     xlab = '5 Minute Interval', ylab = 'Average Number of Steps',
     main = 'Average Weekend Activity Pattern')
```